Viral Sequence Sorting Tools Evaluation

### 2022-11-17 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Refining figures.

```{r}
ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
                   fill=confusion_matrix,
                   color=confusion_matrix)) +
  geom_bar(stat="identity") +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Number of Sequences") +
  ylab("") + 
  facet_wrap(~seqtype, scales = "free") +
  coord_flip()
```
this rule set had the highest precision, but as you can see, this comes with a big sacrifice in recall

@@

@@

>>>>

```{r}
ggplot(confusion_by_taxa, aes(y=count, x=confusion_matrix,
                   fill=confusion_matrix,
                   color=confusion_matrix)) +
  geom_boxplot() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Number of Sequences") +
  ylab("") + 
  facet_wrap(~seqtype, scales = "free")
```
this rule set had the highest precision, but as you can see, this comes with a big sacrifice in recall

@@

## Visualizing the different sets

```{r}
confusion_by_taxa_method <- viruses %>% 
  select(contains("confusion_matrix"), seqtype, Index) %>%
  pivot_longer(cols=contains("confusion_matrix"), 
               names_to="confusion_matrix_type",
               values_to="confusion_matrix_value") %>%
  count(seqtype, Index, confusion_matrix_type, confusion_matrix_value)
```

```{r}
ggplot(confusion_by_taxa_method, aes(y=n, x=confusion_matrix_type,
                   fill=confusion_matrix_value,
                   color=confusion_matrix_value)) +
  geom_boxplot() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Number of Sequences") +
  ylab("") + 
  facet_wrap(~seqtype, scales = "free")
```

```{r}
confusion_by_taxa_method <- viruses %>% 
  select(contains("confusion_matrix"), seqtype, Index) %>%
  pivot_longer(cols=contains("confusion_matrix"), 
               names_to="confusion_matrix_type",
               values_to="confusion_matrix_value") %>%
  count(seqtype, Index, confusion_matrix_type, confusion_matrix_value) %>%
  filter(grepl("true", confusion_matrix_value)) %>%
  mutate(confusion_matrix_type=sub("confusion_matrix_", "", confusion_matrix_type))
```

```{r}
ggplot(confusion_by_taxa_method, aes(y=n, x=confusion_matrix_type,
                   fill=confusion_matrix_value,
                   color=confusion_matrix_value)) +
  geom_boxplot() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=10, angle=90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4))[3:4], 0.5),
                    labels=c( 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4))[3:4], 1),
                    labels=c(
                             "true negative", "true positive")) +
  xlab("Tool Set") +
  ylab("Contig Count") + 
  facet_wrap(~seqtype, scales = "free")
```

@@

```{r}
confusion_by_keep_score <- viruses %>% 
  select(contains("keep_score"), seqtype, Index) %>%
  pivot_longer(cols=contains("keep_score"), 
               names_to="pipeline_type",
               values_to="pipeline_value") %>%
  mutate(pipeline_type=sub("keep_score_", "", pipeline_type)) %>%
  count(seqtype, Index, pipeline_type, pipeline_value)
```

```{r}
confusion_by_keep_score$confusion_matrix <- "true negative"
confusion_by_keep_score$confusion_matrix[confusion_by_keep_score$seqtype=="virus" & confusion_by_keep_score$pipeline_value<1] <- "false negative"
confusion_by_keep_score$confusion_matrix[confusion_by_keep_score$seqtype=="virus" & confusion_by_keep_score$pipeline_value>=1] <- "true positive"
confusion_by_keep_score$confusion_matrix[confusion_by_keep_score$seqtype!="virus" & confusion_by_keep_score$pipeline_value>=1] <- "false positive"
```

```{r}
confusion_by_keep_score$keep_score_visualize <- confusion_by_keep_score$pipeline_value
confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value>1] <- "> 1"
confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==1] <- "1"
confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==0.5] <- "0.5"
confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==0] <- "0"
confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==-0.5] <- "-0.5"
confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==-1] <- "-1"
confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value<=-1] <- "< -1"

confusion_by_keep_score$keep_score_visualize <- factor(confusion_by_keep_score$keep_score_visualize, 
                                       levels=c("< -1", "-1", "-0.5", "0", "0.5","1", "> 1"))
```

```{r}
ggplot(confusion_by_keep_score, aes(x=confusion_matrix, y=n,
                   fill=keep_score_visualize, color=keep_score_visualize)) +
  geom_boxplot() +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16)
  ) +
  scale_color_manual(name = 'Viral Score',
                     values = alpha(c(viridis(6)), 1)) +
  scale_fill_manual(name = 'Viral Score',
                     values = alpha(c(viridis(6)), 0.5)) +
  xlab("Index") +
  ylab("Sequence Count") +
  facet_wrap(~pipeline_type, scales = "free")
